---
title: "May 12 2026- Tidy Tuesday"
output: html_document
date: "2026-05-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidygeocoder)
library(tidyverse)
library(tidytuesdayR)
library(janitor)
library(tinytable)
library(ggplot2)
library(maps)
library(dplyr)
library(circlize)
library(tidyr)

#Download TidyTuesday data
tuesdata <- tidytuesdayR::tt_load("2026-05-12")

#Find datasets for this week
names(tuesdata)

#Name datasets
cities <- tuesdata$cities
links <- tuesdata$links
```

For this dataset, I was interested in Canadian cities with sister cities. I began by filtering the dataset to show cities in the dataset, and using geocode to create additional columns that would show which province/territory each city was in, so that I could look at the density of sister cities by province. 


```{r}
#Check cities in Canada
cdncities <- cities %>%
    filter(country == "Canada")

names(cities)

#Geocode to provinces
cdncities_prov <- cdncities %>%
reverse_geocode(
    lat = lat,
    long = lng,
    method = "osm",
    full_results = TRUE
)

#check work    
head(cdncities_prov)

#rename 'state' to 'province' for clarity in Cdn context
cdncities_prov <- cdncities_prov %>%
  mutate(province = state)

head(cdncities_prov)
```

I then created a map with points for each city, colour coded by province, and used leaflet to create an interactive map that had clickable points to show the city names.

```{r}

#Create map with points for sister cities
ggplot() +
    borders("world", regions = "Canada", fill = "grey") +
    geom_point(data = cdncities_prov,
               aes(x = lng, y = lat, color = province), size = 2, alpha = 0.7) +
    labs(title = "Canadian Cities with Sister Cities by Province (Twinned Cities Dataset)", color = "Province") +
    coord_quickmap() +
    theme_minimal() +
    theme (
        legend.position = "right"
  )


```

```{r}

#Make map with clickable points to show city names 
library(leaflet)
leaflet(cdncities_prov) %>%
  addTiles() %>%
  addCircleMarkers(~lng, ~lat, color = ~factor(province),
  popup = ~name...2, radius = 6)
```


The provinces with the highest density of sister cities appear to be British Columbia and Ontario. I then wanted to look at the sister city links by country. I first tried creating a chord diagram.


```{r}
# join country info for links
links_country <- links %>%
  left_join(cities %>% select(id, country_source = country), by = c("source" = "id")) %>%
  left_join(cities %>% select(id, country_target = country), by = c("target" = "id"))

# make unique country pairs
canada_pairs <- links_country %>%
  filter(country_source == "Canada" | country_target == "Canada") %>%
  mutate(
    country_a = pmin(country_source, country_target),
    country_b = pmax(country_source, country_target)
  ) %>%
  group_by(country_a, country_b) %>%
  summarize(n = n(), .groups = 'drop')

# create matrix
matrix1 <- canada_pairs %>%
  tidyr::pivot_wider(names_from = country_b, values_from = n, values_fill = 0) %>%
  tibble::column_to_rownames('country_a') %>%
  as.matrix()

matrix1 <- matrix1[rowSums(matrix1) > 0, colSums(matrix1) > 0]

country_names <- rownames(matrix1) 
my_colors <- setNames(
  c("red", RColorBrewer::brewer.pal(length(country_names)-1, "Set3")),
  country_names
)
# set 'Canada' to red and the rest unique colors from Set3 palette
my_colors["Canada"] <- "red"

# create chord diagram

circos.clear()
circos.par(gap.degree = 4)
chordDiagram(
  matrix1,
  annotationTrack = "grid",   
  grid.col = my_colors,
  transparency = 0.5,
  preAllocateTracks = list(track.height = 0.08)
)

# Add tilted labels
circos.trackPlotRegion(
  track.index = 1, 
  panel.fun = function(x, y) {
    sector.name = get.cell.meta.data("sector.index")
    xlim = get.cell.meta.data("xlim")
    ylim = get.cell.meta.data("ylim")
    offset <- 1  # move labels further out
    circos.text(
      mean(xlim), ylim[2] + offset, sector.name,
      cex = 0.7, adj = c(0.5, 0), facing = "clockwise", niceFacing = TRUE
    )
  }, 
  bg.border = NA
)
```

I liked this idea, but it was too messy, so I tried again with just the top 10 countries with links to Canada.


```{r}
# Rank countries by the number of links to/from Canada
top_n <- 10  # adjust as needed

# Get counts for top countries 
country_counts <- canada_pairs %>%
  filter(country_a == "Canada" | country_b == "Canada") %>%
  mutate(other_country = ifelse(country_a == "Canada", country_b, country_a)) %>%
  group_by(other_country) %>%
  summarise(n = sum(n)) %>%
  arrange(desc(n))

top_countries <- country_counts$other_country[1:top_n]

# Only keep top countries + Canada, set rest as "Other"
canada_pairs_small <- canada_pairs %>%
  mutate(across(c(country_a, country_b), 
                ~ifelse(!(.) %in% c("Canada", top_countries), "Other", .))) %>%
  group_by(country_a, country_b) %>%
  summarise(n = sum(n), .groups = "drop")

# create new matrix without only top countries & Canada
matrix_small <- canada_pairs_small %>%
  pivot_wider(names_from = country_b, values_from = n, values_fill = 0) %>%
  tibble::column_to_rownames('country_a') %>%
  as.matrix()

# Remove empty rows/cols
matrix_small <- matrix_small[rowSums(matrix_small) > 0, colSums(matrix_small) > 0]

country_names <- rownames(matrix_small)  # or colnames(matrix_small), should be the same set
my_colors <- setNames(
  c("red", RColorBrewer::brewer.pal(length(country_names)-1, "Set3")),
  country_names
)
# set 'Canada' to red and the rest unique colors from Set3 palette
my_colors["Canada"] <- "red"


# create chord diagram
circos.clear()
circos.par(gap.degree = 10, track.margin = c(0.01, 0.01))
chordDiagram(
  matrix_small,
  annotationTrack = "grid",   
  grid.col = my_colors,
  transparency = 0.5,
  preAllocateTracks = list(track.height = 0.08)
)

# Add tilted labels
circos.trackPlotRegion(
  track.index = 1, 
  panel.fun = function(x, y) {
    sector.name = get.cell.meta.data("sector.index")
    xlim = get.cell.meta.data("xlim")
    ylim = get.cell.meta.data("ylim")
    offset <- 1  # move labels further out
    circos.text(
      mean(xlim), ylim[2] + offset, sector.name,
      cex = 0.7, adj = c(0.5, 0), facing = "clockwise", niceFacing = TRUE
    )
  }, 
  bg.border = NA
)
```
