---
title: "2026-05-19"
output: html_document
date: "2026-05-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(janitor)
library(tinytable)
library(tidyr)

#download TidyTuesday data
tuesdata <- tidytuesdayR::tt_load('2026-05-19')

#name datasets
member_participation_stats_by_country <- tuesdata$member_participation_stats_by_country
metadata_coverage_stats_by_country <- tuesdata$metadata_coverage_stats_by_country

```

Firstly, I filtered the member participation stats to view Canadian data. I then created a plot to show the total members over time in Canada, current up to 2026.

```{r}
#filter dataset to show Canadian data
cdn_member_stats <- member_participation_stats_by_country %>%
    filter(iso3_code == "CAN")

colnames(cdn_member_stats)
```

```{r}
#create plot of total members over time in Canada
ggplot(cdn_member_stats, aes(x = current_up_to, y = total_members)) +
  geom_line() +
  labs(title = "Trend of Total Members Over Time in Canada")
```

We can see that the total number of members in Canada has increased over the last ~6 years. I then wanted to look at some metadata coverage stats across countries. I chose to use the G7 countries to create graphs with enough countries for some comparison, but not so many that the graph became jumbled/messy. 

```{r}
#create graph of ORCID coverage over time
#filter by G7 countries & journal article document type
    timeseries_orcid <- metadata_coverage_stats_by_country %>%
      filter(document_type == "journal-article",
             iso3_code %in% c("CAN", "FRA", "DEU", "ITA", "JPN", "GBR", "USA")) %>%
      group_by(current_up_to, iso3_code) %>%
      summarise(orcid_coverage = sum(with_orcid, na.rm=TRUE)/sum(n_dois, na.rm=TRUE))

    ggplot(timeseries_orcid, aes(x = as.Date(current_up_to), y = orcid_coverage, color = iso3_code, size = iso3_code == "CAN")) +
      geom_line() +
      scale_size_manual(values=c(1,2), guide=FALSE) +
      labs(title = "ORCID Coverage Over Time For G7 Countries (Journal Articles)", x = "Date", y = "% with ORCID")

#create graph of funding acknowledgements 
   timeseries_funding <- metadata_coverage_stats_by_country %>%
      filter(document_type == "journal-article",
             iso3_code %in% c("CAN", "FRA", "DEU", "ITA", "JPN", "GBR", "USA")) %>%
      group_by(current_up_to, iso3_code) %>%
      summarise(coverage_funding = sum(acknowledges_funding, na.rm=TRUE)/sum(n_dois, na.rm=TRUE))

    ggplot(timeseries_funding, aes(x = as.Date(current_up_to), y = coverage_funding, color = iso3_code, size = iso3_code == "CAN")) +
      geom_line() +
      scale_size_manual(values=c(1,2), guide=FALSE) +
      labs(title = "Funding Acknowledgement Coverage Over Time For G7 Countries (Journal Articles)", x = "Date", y = "% with Funding Acknowledgment")
    
```